### 2019-02-10 paper.tex
Wip

\usepackage{microtype}
\usepackage{graphicx}
\usepackage{subfigure}
\usepackage{booktabs}
\usepackage{hyperref}
\usepackage{listings}

\lstset{
  aboveskip=12pt,
  belowskip=0pt,
  basicstyle=\fontsize{8.5pt}{8.5pt}\ttfamily,
  breaklines=true,
  captionpos=b,
  frame=tb,
  framexbottommargin=5pt,
  framextopmargin=5pt,
}

\usepackage{enumitem}

\setlist[description]{
  itemsep=0pt,
  topsep=0pt,
  labelindent=1em,
  leftmargin=1em,
}

\setlist[enumerate]{
  itemsep=0pt,
  topsep=0pt,
}

\setlist[itemize]{
  itemsep=0pt,
  topsep=0pt,
  labelindent=1em,
  leftmargin=1em,
}

% Use the following line for the initial blind version submitted for review:
\usepackage{sysml2019}

% If accepted, instead use the following line for the camera-ready submission:
%\usepackage[accepted]{sysml2019}

% The \sysmltitle you define below is probably too long as a header.
% Therefore, a short form for the running title is supplied here:
%\sysmltitlerunning{Guild AI: NEEDS A TITLE}

@@

\twocolumn[
\sysmltitle{Guild AI: Package management for machine learning models}

@@

\begin{sysmlauthorlist}
\sysmlauthor{Garrett Smith}{guild}
\end{sysmlauthorlist}

\sysmlaffiliation{guild}{Guild AI, Chicago, Illinois, USA}

@@

% You may provide any keywords that you
% find helpful for describing your paper; these are used to populate
% the "keywords" metadata in the PDF but will not be shown in the document
\sysmlkeywords{Guild AI, SysML}

@@

\begin{abstract}

  We present Guild AI, an open source toolkit that facilitates model
  reuse for application development by applying traditional software
  packaging constructs to the domain of machine learning. Package
  management strategies are central to successful software
  ecosystems. Examples include npm for JavaScript \cite{npm}, APT for
  Debian \cite{apt}, and pip for Python \cite{pip}. Such tools excel
  at creating traditional software packages. Machine learning
  applications however present unique requirements that call for tool
  specialization. Guild AI combines the proven effectiveness of
  package managers with novel features to enable machine learning
  model reuse for application development across a variety of use
  cases.

\end{abstract}
]

@@

\printAffiliationsAndNotice{}  % leave blank if no need to mention equal contribution
%\printAffiliationsAndNotice{\sysmlEqualContribution} % otherwise use the standard text.

\section{Introduction}

Code reuse is a central tenet in software development and is supported
by myriad techniques, abstractions, and tools. Effective code reuse
allows developers to leverage prior work in a consistent, reliable way
that saves time and improves software quality. In this paper we
consider code reuse in terms of \emph{packaged software} in the
context of machine learning.

\subsection{Traditional Packaged Software}

Packaged software is software that is organized in a way that can be
easily installed and used by \emph{users}. A user is a consumer of
software, rather than a producer or developer. A user may be software
developer, in the case of consuming software libraries or frameworks,
or she may be an end-user who uses software as a program with a
graphical or command line interface. In both cases, packaged software
may be provided in units that the user can install and start using
quickly and reliably.

This is an example of installing packaged software:

{\footnotesize
\begin{verbatim}
 $ pip install tensorflow
\end{verbatim}}

By running this command, a user is using a \emph{package manager}---in
this case the program \texttt|pip|---to install a package named
\verb|tensorflow|. With this innocuous command, the user implicitly expects
a number of things:

\begin{itemize}
\item The Python module \verb|tensorflow| should be installed.
\item Additional Python modules required by \verb|tensorflow| should
  also be installed.
\item Installed software should work as expected on the target system.
\item At a future date \verb|tensorflow| may be similarly uninstalled
  along with its no-longer-used required packages.
\item The act of installing and uninstalling \verb|tensorflow| should
  not break or otherwise destabilize other installed software on the
  target system.
\end{itemize}

Such package management facilities are essential to code reuse at the
system level. Package managers allow users to maintain complex
software systems with simple commands in a way that preserves their
integrity. They support rapid experimentation with new software as
users are free to install, use, and evaluate packages with the
confidence that they can safely remove that software later.

\iffalse
Package managers promote healthy software ecosystems by facilitating
software \emph{publishing}. Software authors use packaging tools to
create the installable software units---\emph{packages}---that are so
easily consumed by users. By publishing packaged software, developers
signify that their software will work as advertised if installed using
the applicable tools.

Software ecosystems that are enabled by package managers include
Python, JavaScript, Ruby, Go, and R. The GNU/Linux operating system
itself is an ecosystem of package management ecosystems including
packaging schemes from Debian, RedHat, Gentoo, and Arch
distributions. Package managers are established cornerstones of
software reuse across languages and operating systems.
\fi

\subsection{Code Reuse and Machine Learning Models}

A \emph{machine learning model} is challenging to define as models
have different contexts. In the most general context, models are
mathematical concepts or language that represents a system. In the
context of machine learning, models may be represented by computer
source code, either explicitly as declared structures, or implicitly
by way of executable instructions that perform model related tasks. A
\emph{trained model} is a model that has learned some representation
of a system through an automated process of optimization.

For the purpose of our discussion, a machine learning model is a
software representation of a specific mathematical concept that can be
trained by a computer. Models may be in various states:
\emph{uninitialized}, \emph{initialized}, \emph{partially trained},
\emph{fully trained}, etc. Regardless of state, the distinguishing
feature of machine learning models for our purpose is that they
support runnable operations.

Models---and in particular \emph{trained models}---are similar to
traditional software in that they can be executed to perform
computations. In fact, the purpose of training a model is to create a
computer program. The difference between a trained model and a human
authored program is one of method: a trained model is learned by an
algorithm while a human authored program is written manually.

\iffalse
While models may be represented in part by traditional installed
software---e.g. an imperative TensorFlow program that defines a model
graph structure when executed---models require more to be construed as
reusable code: \emph{models must support operations on novel data}. If
they cannot be applied to novel data---either for training or for
inference---they cannot be considered reusable software.
\fi

The remainder of this paper will outline an approach to package
management that incorporates \emph{model operations} and \emph{run
  management} as an enabler of code reuse for machine learning models.

@@

\subsubsection{Model zoos}

A common approach to sharing and reusing machine learning models is a
\emph{model zoo}---an online repository of reusable models that
includes source code, examples, documentation, and in some cases
pretrained models. The Caffe Model Zoo \cite{caffezoo} was one of the
first model zoos and has inspired several others \cite{tfzoo,mxzoo}.

Model zoos encourage code reuse by consolidating related
models---e.g. based on underlying library implementation or
application domain---and providing consistent documentation and
repeatable methods for model use.

Model zoos, however, do not typically provide tools for packaging,
distributing, and using models beyond those of publishing source code
and documentation (e.g. GitHub, web sites, etc.)

\subsubsection{TensorFlow Hub and Keras Applications}

TensorFlow Hub modules \cite{tfhub} and Keras Applications
\cite{kerasapps} are examples of framework level facilities that
provide a higher degree of model reuse vis-\'a-vis model zoos. Both
TensorFlow Hub modules and Keras application support tools and
patterns for packaging, distributing, and using model source code and
associated pretrained models. They emphasize the use of \emph{transfer
  learning} for applying pretrained models to data from different
domains.

\subsubsection{Software Packages and Source Code Repositories}

Models are commonly reused through traditional software distribution
channels: \emph{packages} and \emph{source code repositories}. In both
cases, model source code is provided along with varying levels of
documentation and supporting tools. Unlike model zoos and framework
facilities, the level of code reuse available through these channels
is spotty and inconsistent.

@@

Guild AI combines elements from model zoos, framework packages, and
traditional software channels to provide a state of the art facility
for packaging, distributing and using machine learning models.

As with model zoos, users can discover Guild AI models, including
documentation. Guild AI additionally supports installation and direct
use of models through operations.

Guild AI superficially resembles TensorFlow Hub in that users can
discover, install, and use models quickly and easily. Guild AI however
differs considerably in its implementation. Guild AI does not use a
programming interface but instead supports its features by way of
declared configuration and system level tools. Guild AI packaging
facilities can be applied to any machine learning model
implementation.

Guild AI also supports the process of \emph{package maintenance},
where a third party---a package maintainer---creates usable packages
without necessarily coordinating with upstream model developers. This
is discussed in Section \ref{sec:package-maintainer}.

As with traditional software packages and source code repositories,
Guild AI supports a federated approach to software distribution. Users
are free to use Guild AI to create and distribute reusable models in a
variety of ways. Refer to Section \ref{sec:use-cases} for coverage of
various use case.

\section{Concepts}

This paper relies on a number of core concepts, which are defined
below.

\subsection{Projects}

A \emph{project} is a directory that contains model related
code. Model \emph{developers} maintain projects for their
models. Model \emph{users} maintain projects for their applications,
which may use models for training, inference, etc.

Projects are central to model development and use. Guild AI supports
project based workflow as described in Section
\ref{sec:streamline-workflow}.

\subsection{Guild File}

A \emph{Guild file} is a plain text file containing Guild AI related
configuration. The file is named \verb|guild.yml| and is typically
located in the root directory of a Guild AI enabled project.

Listing~\ref{lst:guild-file} illustrates a simple Guild file.

\begin{figure}
\begin{lstlisting}[
    caption=Simple Guild file example,
    label={lst:guild-file}]
model: mnist
description:
  Logistic regression classifier
  of MNIST digits
operations:
  train:
    main: train_mnist
    flags:
      batch-size: 32
      train-steps: 1000
  evaluate:
    main: eval_mnist
\end{lstlisting}
\end{figure}

\subsection{Packages}

A Guild AI \emph{package} is a Python wheel distribution \cite{wheels}
that was generated using the \verb|guild package| command. Guild AI
packages are generated using \verb|setuptools|. They contain Guild AI
specific metadata that advertises models they contain.

\subsection{Models}

A Guild AI \emph{model} is a representation of a machine learning
model that is defined in a Guild file. Models define \emph{operations}
(Section \ref{sec:operations}) that perform model related tasks such
as training or evaluating as well as \emph{resources} that operations
may require (Section \ref{sec:resources}).

Models may be defined according to the underlying architecture or
intended use of the model.

\subsection{Operations}
\label{sec:operations}

An \emph{operation} is an action that can be performed on a
model. Operations are part of a model definition.

Guild AI does not prescribe any particular set of operations that must
be defined for a model. Model developers are free to name operations
as they see fit, though Guild AI does encourage the use of consistent
naming conventions.

Below are some common operation names.

\begin{description}
\item[prepare] prepare a dataset for training
\item[train] train a model from scratch
\item[transfer-learn] train a model using transfer learning
\item[finetune] finetune a trained model
\item[evaluate] evaluate/test a model
\item[classify] use a trained model for classification
\item[compress] compress a trained model (e.g. using quantization,
  etc.)
\item[export-and-freeze] export a frozen inference
  graph\footnote{TensorFlow specific}
\item[tflite] generate a TF Lite file from a frozen
  model\footnotemark[\value{footnote}]
\end{description}

This list illustrates that model operations can be used to perform
tasks beyond just training and evaluation. Feature rich models can
encompass all of the operations associated with a machine learning
development pipeline, from data acquisition and preparation through
training and optimization to application deployment and monitoring.

\subsection{Flags}

\emph{Flags} are parameters that can be set or modified by users when
they run an operation. Flags are used as inputs to the underlying code
that implements an operation.

Below is an example of using a flag to set the \emph{learning rate}
for a \emph{train} operation.

{\footnotesize
\begin{verbatim}
 $ guild run train learning-rate=0.0001
\end{verbatim}}

\subsection{Resources}
\label{sec:resources}

A Guild AI \emph{resource} represents a set of source files that can
be made available as inputs to a model operation. Resources play an
important role in Guild AI's ability to effective use
models. Resources can be used to resolve project files, download files
from URLs, unpack archives, and reference files generated by other
operation.

Listing~\ref{lst:resource} contains a resource definition for a
model. The resource \verb|idx-data| defines four source files that
will be made available to any operation that requires that resource.

\begin{figure}
\begin{lstlisting}[
    caption=Model resource example,
    label={lst:resource}]
model: mnist
resources:
  idx-data:
    description:
      MNIST dataset in compressed
      IDX format
    sources:
      - train-images-idx3-ubyte.gz
      - train-labels-idx1-ubyte.gz
      - t10k-images-idx3-ubyte.gz
      - t10k-labels-idx1-ubyte.gz
\end{lstlisting}
\end{figure}

\subsection{Runs}

A \emph{run} is a file system artifact---specifically a directory
created and managed by Guild AI---that contains metadata and files
generated by an operation. A run may also refer to the operation
system process associated with a running operation.

Runs are further described in Section \ref{sec:runs}.

\subsection{Run Directories}
\label{sec:run-dir}

A \emph{run directory} is a directory used for a run. When an
operation is run, the run directory is used as the operating system
process working directory. Operations should write run artifacts
within the working directory to ensure they are associated with the
run.

By default, Guild AI generates a new directory for each run and
creates links to required sources within that directory. Once the
directory is prepared, Guild AI starts the run operating system
process, which may read sources within the run directory.

\subsection{Environments}

Guild AI supports \emph{environments}, which are isolated runtime
environments used for installing and using models. Guild AI supports
environments created using
\href{https://virtualenv.pypa.io}{\emph{virtualenv}}.

\subsection{Workflows}

Guild AI \emph{workflows} are patterns of model use that are enabled
by running operations in specific order. Worflows are not currently
formal constructs in Guild AI but are used by convention and may be
implemented by other programs.

\section{Use Cases}
\label{sec:use-cases}

The package management facilities in Guild AI enable a number of use
cases, which are described below. In each case, the following roles
may be referenced:

\begin{description}
\item[user] person using a model developed by someone else
\item[developer] person developing a model, either from scratch or
  adapting the work of others
\item[packager] person neither using nor developing a model but who
  packages models created by developers for users
\end{description}

\subsection{Publish a Model for Use}
\label{sec:publish-model}

A \emph{developer} may encourage others to use her model by
documenting the steps needed to install, adapt, and use it. She may be
motivated to simplify the user experience so that her work is more
easily and successfully applied. The benefit to her may be quality
feedback from expert users, improved reputation, a sense of
contribution, or her colleagues may simply expect that her work be
documented and easy to use.

While documenting a model is undoubtedly helpful, the process can be
burdensome for the developer. Similarly, following documented steps
can be burdensome for users. Guild AI can be used to simultaneously
document model features and support model usage through operations.

Consider these instructions for model use:

\setlength{\parindent}{1em}
\begin{enumerate}
\item Install required packages
\item Download the data
\item Prepare the data for training
\item Train the model
\item Export the trained model for deployment
\end{enumerate}
\setlength{\parindent}{0em}

While these steps provide invaluable information, they still rely on
the user to carefully execute them and troubleshoot issues. The
process takes time and is subject to error at any step.

Guild AI can be used to both document and automate these
steps. Listing~\ref{lst:published-model} illustrates a Guild file that
might be used to codify the steps as model operations.

The operations can be run using the Guild AI command line interface as
follows:

{\footnotesize
\begin{verbatim}
 $ guild run prepare-data
 $ guild run train
 $ guild run export-model
\end{verbatim}}

Guild AI additionally provides model help, detailing available
operations, supported flags, and operation dependencies.

A user can get help for the model by running:

{\footnotesize
\begin{verbatim}
 $ guild help
\end{verbatim}}

Documentation is automatically generated from the model description,
making Guild AI models self documenting.

The developer can chose to distribute her model as a
project---e.g. making it available as a GitHub repository---or she may
package the model and deploy it as an installable Python wheel
distribution.

\begin{figure}
\begin{lstlisting}[
    caption=Sample project Guild file,
    label={lst:published-model}]
model: sample
operations:
  prepare-data:
    description:
      Prepare data for training
    main: prepare
    requires: data
  train:
    description: Train the model
    main: train
    requires: prepared-data
  export-model:
    description:
      Export the model for deployment
    main: export_model
    requires: trained-model
resources:
  data:
    description: Data for training
    sources:
      - url: http://my.co/data.tgz
  prepare-data:
    description: Prepared data
    sources:
      - operation: prepare-data
  trained-model:
    description: Trained model
    sources:
      - operation: train
\end{lstlisting}
\end{figure}

\subsection{Use a Published Model}
\label{sec:use-model}

The flip side of the use case outlined in
Section~\ref{sec:publish-model} is from the \emph{user} point of
view. A user is interested in taking the work of a model developer and
applying it, either by adapting the model to a new application or by
integrating it as-is into a project.

Users commonly rely on documentation to effectively use models. If a
project is poorly documented or otherwise hard to work with, the user
may choose not to use it. This would be a lost opportunity for both
user and developer and would hinder model adoption and improvement.

As described in Section~\ref{sec:publish-model}, a published model can
be used by running its operations. From the standpoint of the user,
the model can be consumed either by cloning the project
repository---e.g. as in the case of GitHub published models---or by
installing the model as a Python package. If, for example, the project
was published as the package \verb|sample-project|, a could user run
the commands below to quickly make use of its features.

\setlength{\parindent}{0.67em}

{\footnotesize\emph{Install the package}}

{\footnotesize
\begin{verbatim}
 $ guild install sample-project
\end{verbatim}}

{\footnotesize\emph{View package help}}

{\footnotesize
\begin{verbatim}
 $ guild help sample-package
\end{verbatim}}

{\footnotesize\emph{Use the model}}

{\footnotesize
\begin{verbatim}
 $ guild run sample:prepare-data
 $ guild run sample:train
 $ guild run sample:export-model
\end{verbatim}}

\setlength{\parindent}{0em}

This example highlights the usefulness of packaging models for
reuse. Packaged models in this case consist of more than the
underlying model code---they include essential configuration that lets
Guild AI \emph{execute model operations}.

\subsection{Package Third Party Models}
\label{sec:package-maintainer}

Complex software ecosystems such as GNU/Linux have formalized the role
of a software \emph{packager}---or \emph{package maintainer}. The goal
of a packager is to improve the value of a software distribution by
creating high quality, useful packages that are easy to install and
use. Packagers are often independent of the upstream software projects
they package, serving as intermediaries between the software
developers and distribution users.

Guild AI supports this pattern by allowing packagers to incorporate
upstream model sources into Guild packages, patching them as needed to
provide the desired functionality to users.

Consider a case where a model developer is overloaded with research
tasks and doesn't have time to create model documentation, much less
publish models as described in Section~\ref{sec:publish-model}. An
independent \emph{packager} may step up to create a package instead.

A packager must assume that upstream sources are available \emph{as
  is} and cannot expect developers to fix bugs or add new
features---though he may certainly submit requests and remain
hopeful. Packagers therefore must be prepared to \emph{patch} model
code as needed to get it to work as expected.

\iffalse
Guild AI provides various means of patching model source to support
packagers in this role. Section~\ref{sec:patching} describes patching
methods.
\fi

\subsection{Team Collaboration}

The roles of \emph{developer}, \emph{user}, and \emph{packager} often
exist within a single organization or team. Guild AI supports team
collaboration by supporting each role as a separate concern.

Within a typical organization structure, the may roles map as follows:

\setlength{\parindent}{1em}
developer $\rightarrow$ \emph{research scientist}

user $\rightarrow$ \emph{software engineer}

packager $\rightarrow$ \emph{research engineer}
\setlength{\parindent}{0em}

Of course this is a rough mapping and not all organizations use this
structure. Whatever the job titles and division of labor, Guild AI
supports the three facets of collaboration: model development, model
use, and intermediation between developer and user as needed to
improve usability without unduly burdening developers.

\subsection{Streamline Model Developer Workflow}
\label{sec:streamline-workflow}

In the case of a single developer who has no interest in publishing
models or otherwise supporting users, Guild AI model abstractions
provide value. By implementing models as units of operations, as is
the case with Guild AI model publishing, developers ensure their
models are easy to use from the early stages of development. This
streamlines the development process as developers can repeatedly run
and iterate through model operations using high level commands with
well defined, self documenting interfaces.

\iffalse
\subsection{Manage Run Artifacts}

The final use case discussed involves managing runs and tracking run
artifacts. During both model development and use, it's helpful to
formally manage run artifacts both generated and consumed by model
operation. This is similar to source code revision control, where
developers can move backward and forward in time to view their
work. In the case of Guild AI, the revisions are the results of data
preparation, training experiments, model optimizations---anything that
might be generated by an operation.

Section~\ref{sec:runs} describes Guild AI run management.
\fi

\section{Using Models}

Section~\ref{sec:use-model} describes the case of using published
models. This section provides more details on model usage in Guild AI.

\subsection{Packaged Models}

Packaged models are models that have been bundled with supporting code
and metadata into a Python wheel distribution. To generate a package,
a developer or packager must include a \verb|package| definition in
the project Guild file and then run the \verb|guild package| command.

Listing~\ref{lst:package} contains a sample \verb|package| definition.

\begin{figure}
\begin{lstlisting}[
    caption=Sample package definition,
    label={lst:package}]
package: sample-project
version: 1.0
description: Sample packaged project
url: https://my.co/packages/sample
author: John Doe
author-email: john@my.co
license: Apache 2.0
requires: [keras, Pillow, matlab]
\end{lstlisting}
\end{figure}

Given a Guild file with a package definition, developers can generate
a Python wheel distribution, which in turn can be used to install
package models. Developers can additionally upload packages to PyPI,
making them available to discover and install.

Guild AI models are installed like any other Python package. Users may
use Guild AI to install the package:

{\footnotesize
\begin{verbatim}
 $ guild install sample-project
\end{verbatim}}

Or alternatively use pip:

{\footnotesize
\begin{verbatim}
 $ pip install sample-project
\end{verbatim}}

Guild AI package requirements may be included in requirements files
(e.g. \verb|requirements.txt|) or as dependencies of other Python
package distributions.

Package models are convenient for using models when a user doesn't
need to modify the model source code. To use models and also have the
option of modifying model source code, a user should use project
models (Section \ref{sec:project-models}).

\subsection{Project Models}
\label{sec:project-models}

Project models are defined in a project Guild file---i.e. a file named
\verb|guild.yml| that resides in the top level of a project
directory. Users may run operations on project models by running them
from within the project directory or by referring the project
directory in the operation specified for the \verb|guild run| command.

Unlike packaged models, project models do not need to be installed
using a command---they are available from the project itself. A user
need only change to the project directory to run model operations.

\subsection{Listing Available Models and Operations}

A user may list available models using the \verb|guild models|
command. This command lists packaged models and, if the command is run
within a project, project models.

Each model has one or more operations, which can be listed using the
\verb|guild ops| command.

The ability to discover available models and operations is a key
feature enabling model use in Guild AI. Without formally defining
these, users must infer model capabilities from documentation and
source code.

\subsection{Getting Help for Models}

A user may run the \verb|guild help| command to show details about a
model, its operations, operation flags, and resource
dependencies. This is similar to the \emph{man} facility on many
systems.

A user may also get help for a particular operation by using the
\verb|--help-op| option with the \verb|guild run| command. This is
similar to the \verb|--help| option available for many programs.

Figure~\ref{fig:op-help} contains sample help for the \verb|train|
operation of the \verb|gpkg.mnist/cnn| package model.

\begin{figure}
\begin{lstlisting}
Usage: guild run [OPTIONS] cnn:train
[FLAG]...

Train the CNN

Use 'guild run --help' for a list of
options.

Dependencies:
  mnist-dataset  Yann Lecun's MNIST
                 dataset in compressed
                 IDX format

Flags:
  batch-size  Number of images to
              include in a training
              batch (100)
  epochs      Number of epochs to train
              (10)
\end{lstlisting}
\caption{Sample operation help}
\label{fig:op-help}
\end{figure}

It is important to note that model and operation help is automatically
generated from model descriptions. This facility lowers the cost of
model documentation for developers and gives users a consistent
interface for discovering model capabilities and the commands for
using them.

\subsection{Running Operations}

Once a model is available---either as an installed package model or
defined in a project---it can be used by running \emph{operations}. An
operation is an action that can be performed on a model. Operations
both \emph{define and implement} model capabilities. The process of
using a model is the process of running operation.

Operations may define an entire application development life cycle for
one model or a host of models that work together.

Section~\ref{sec:operations} lists common operation names.

The act of running an operation starts a \emph{run}, which is both an
operating process and a Guild AI managed file system artifact that
contains run details and generated output.

A user can run an operation using the \verb|guild run| command.

Section~\ref{sec:runs} describes how runs are managed.

\subsection{Viewing and Comparing Runs}

As users and developers run model operations, they can view the
results in various ways. Guild AI provides a comprehensive command
line interface for managing runs, which is outlined in
Section~\ref{sec:runs}.

Guild AI also provides various methods for visualizing runs. Guild
View is an application provided with Guild AI that displays run
details including run files. Guild AI integrates TensorBoard \cite{tb}
as a visualization tool for viewing TensorFlow event logs generated by
model operations.\footnote{Not all operations generate TensorFlow
  event logs---these must be explicitly written during a run by the
  operation.}

\subsection{Accessing Generated Artifacts}

Operations generate \emph{artifacts}, which are Guild AI managed files
stored on the local file system. Guild AI does not attempt to
obfuscate generated artifacts---they are maintained in the run
directory as written during the operation.

Artifacts are routinely used as inputs to subsequent operations. For
example, a TensorFlow checkpoint generated by a \verb|train| operation
may be used as input to an \verb|export-and-freeze| operation. The
process of resolving required operation sources is automatic and part
of Guild AI's run support.

Artifacts may also be accessed directly by users given their file
system paths. A user can list the full paths for all files associated
with a run using the \verb|guild runs info| command with the
\verb|-ff| option. Using their paths, a user can copy desired files or
link to them as needed.

Runs may also be \emph{exported} to a location using the
\verb|guild export| command. Exported runs are copied---or moved, if
specified---to a specified file system directory. This can be used to
create backups or to transfer runs to another location.

\subsection{Isolating Runs Using Environments}

The discussion of using models to this point has not mentioned
\emph{environments}---environment are optional for using Guild
AI. Environments are useful, however, for isolating installed packages
and runs on a system.

When running Guild AI operations in an activated environment, the
following are environment specific:

\begin{itemize}
\item Installed packages and therefore installed models
\item Runs
\end{itemize}

Environments are created using the \verb|guild init| command. Once
created, an environment must be explicitly activated using the
\verb|source guild-env| command.

Environments are useful isolating project level work. It is common to
have a separate environment for each project or work stream.

\section{Managing Runs}
\label{sec:runs}

A \emph{run} is a formally managed file system artifact generated by
an operation. Runs may also refer to the operating system process
associated with a running operation.

Runs contain critical information for tracking and understanding what
happened during an operation:

\begin{itemize}
\item Run metadata such as the associated model and operation, status,
  start and stop times, operating system command and environment, and
  resolved dependencies
\item Output generated by the run---specifically the process standard
  output and standard error
\item Files used by the operation (resolved resource sources)
\item Files generated by the operation
\end{itemize}

\subsection{Starting a Run}

Runs are started using the \verb|guild run| command, referencing a
model operation. Operations may be run with additional flag values
provided in the format \verb|NAME=VALUE|.

Here's an example of running an operation with flag values:

{\footnotesize
\begin{verbatim}
 $ guild run mnist:train \
     steps=10000 \
     batch-size=64 \
     learning-rate=0.0001
\end{verbatim}}

\subsection{Listing Runs}

Runs may be listed using the \verb|guild runs| command. Runs are
displayed with their unique ID, operation name, start time, status,
and user label.

Listed runs may be filtered by status, operation, and label.

\subsection{Getting Run Information}

Information about one or more runs may be displayed using the
\verb|guild runs info| command. By default, high level run information
such as the run ID, operation, start time, command, exit status, and
operating system pid are displayed. Additional command options may be
used to view more run information including files, flag values,
dependencies, and operation output.

\subsection{Deleting Runs}

One or more runs may be deleted using the \verb|guild runs rm|
command. Runs may be deleted using status, operation, or label.

Guild AI will restore deleted runs with \verb|guild runs restore| and
permanently delete runs using \verb|guild runs purge|.

\section{Developing Models}

As discussed above, published models---deployed from installed
packages or as projects---are used by running
\emph{operations}. Operations generate \emph{artifacts} that can be
used in subsequent operations or used as final outputs in an
application development workflow.

In this section we discuss how models are created using Guild AI.

\subsection{Developing a Model from Scratch}

When developing a model from scratch, a developer may ask two
questions:

\begin{itemize}
\item What is the defining characteristic of the model?
\item What can users do with the model?
\end{itemize}

The answers to these questions will inform the decision of what to
name the model and its operations. These values can be stubbed out in
a Guild file to provide a scaffold for implementation.

In some cases, this may simply be a matter of selecting a model name
and creating a \verb|train| operation. Listing~\ref{lst:pet-stub}
illustrates a straightforward pet classifier.

\begin{figure}
\begin{lstlisting}[
    caption=Sample Guild file for a pet classifier,
    label={lst:pet-stub}]
model: pets
description:
  Pet classifier using a CNN
operations:
  train:
    description: Train the classifier
    main: train
    flags:
      images:
        description:
          Directory containing the pet
          images to train
        required: yes
      epochs:
        description:
          Number of epochs to train
        default: 100
\end{lstlisting}
\end{figure}

With even a simple stub, a developer can formalize the model and what
it does. This stage clarifies the model interface before the
functionality is implemented.

With a basic skeleton in place, the developer can \emph{implement}
model operations. In the case of the pet classifier in
Listing~\ref{lst:pet-stub} this means implementing train support in a
file named \verb|train.py| that resides in the project directory.

Operations can be developed and tested early using Guild AI. Each time
\verb|guild run| is used to test an operation, Guild AI generates a
distinct, tracked run. Runs may be compared over time to help answer
questions or resolve issues.

Guild AI provides two additional mechanisms to help model developers
implement operations. Both effect the \emph{run directory}
(Section~\ref{sec:run-dir}) used by Guild AI.

The first mechanism is the use of \verb|--run-dir| with the
\verb|guild run| command. This option specifies an alternative
directory for the run. This is helpful when the developer doesn't want
to generate new runs during the development process.

The second mechanism is the use of \verb|--stage| with the
\verb|guild run| command. This option may be used to stage a run in an
alternative directory without running the operation. This is helpful
when the developer wants to study the run directory layout or to run
low level commands or scripts during the development process.

\iffalse
The developer can continue to evolve the model operations as needed to
satisfy user requirements. This may include new operations or
additional operation flags, which give users more flexibility when
running an operation. The developer may similarly add new models to
the Guild file.
\fi

\subsection{Adapting Existing Model Code}

In many cases, a developer will already have working model code she
wants to incorporate into a formalized model definition. As with the
role of \emph{packager} (Section~\ref{sec:package-maintainer}), a
developer may be required to work with upstream sources she cannot
easily change. Guild AI provides a number of features that support the
adaptation of existing model code into usable models.

The operation \verb|main| attribute specifies a Python main module to
run. In many cases, a developer can simply reference an existing
Python script to incorporate it into an operation. In cases where a
script doesn't exist---for example, the model code is in a Jupyter
Notebook---the developer must create a new script that contains the
applicable model code and specify it for the operation.

Operation \emph{flags} specified in the form \verb|NAME=VALUE| as
arguments to the \verb|guild run| command are passed to the Python
main module as command line arguments in the form
\verb|--NAME VALUE|. For example, assuming an operation \verb|train|
with a main module \verb|train|, the command
\verb|guild run train epochs=100| would be translated to the Python
command \verb|python -m train --epochs 100|. In this way developers
can parameterize scripts using operation flags.\footnote{It is
  standard practice to use Python's argparse module to support named
  command line options following the getopt convention \cite{getopt}.}

\iffalse
If a flag name doesn't correspond to the option name expected
(e.g. the train script expected the option \verb|num_epochs|) the
developer can specify an alternative argument name for a flag using
the \verb|arg-name| attribute.
\fi

\subsection{Patching}
\label{sec:patching}

In cases where a developer cannot modify upstream sources or chooses
not to (e.g. to simplify ongoing maintenance) he may choose to patch
model sources for use in a model. Guild AI supports two methods of
patching.

The first method is to run a post processing script on required model
sources. This can be specified using a \verb|post-process| command for
a model source that applies patch files to upstream sources. Guild AI
will use the patched versions for related operations.

The second method is to wrap upstream scripts and monkey patch the
Python environment as needed to change behavior. An example of this
can be found in the \verb|gpkg.slim| package, which patches the
TensorFlow \verb|slim.losses.softmax_cross_entropy| module to add
support for balanced weight training. This technique successfully adds
a useful feature while avoiding the need to modify the upstream train
support in TF Slim.\footnote{While one could argue that it would be
  better to submit a pull request with the new functionality to
  upstream authors, there are always cases where this is not an
  option. Patching in these cases may be the only way to fix issues or
  add features.}

\subsection{Model Definition Reuse}

Guild AI provides a number of features that support model definition
reuse.

\subsubsection{Configuration Inheritance}

Guild files may contain \verb|config| sections that that can be
inherited by models definitions. Listing~\ref{lst:inheritance}
contains two models, each inheriting the properties of shared
configuration. The shared config \verb|model-base| defines the common
model operations and uses \emph{parameters} to define a model name and
default learning rate.

\begin{figure}
\begin{lstlisting}[
    caption=Configuration inheritance,
    label={lst:inheritance}]
- config: model-base
  operations:
    train:
      main: train --model {{model-name}}
      flags:
        epochs: 100
        learning-rate: {{default-lr}}

- model: logreg
  extends: model-base
  params:
    model-name: logreg
    default-lr: 0.01

- model: cnn
  extends: model-base
  params:
    model-name: cnn
    default-lr: 0.0001
\end{lstlisting}
\end{figure}

Configuration inheritance can be applied across Guild files---e.g. to
configuration defined in packages or other projects. This is useful
for extending model definitions from other
developers. Listing~\ref{lst:extend-slim} demonstrates how a model can
inherit all of the operations of another model with a single line of
configuration.

\begin{figure}
\begin{lstlisting}[
    caption=Extending PNASNet Mobile,
    label={lst:extend-slim}]
- model: my-classifier
  extends: gpkg.slim.models/pnasnet-mobile
\end{lstlisting}
\end{figure}

>>>>

\usepackage{microtype}
\usepackage{graphicx}
\usepackage{subfigure}
\usepackage{booktabs}

\usepackage{hyperref}

\newcommand{\theHalgorithm}{\arabic{algorithm}}

% Use the following line for the initial blind version submitted for review:
%\usepackage{sysml2019}

% If accepted, instead use the following line for the camera-ready submission:
\usepackage[accepted]{sysml2019}

\sysmltitlerunning{Guild AI: what to put?}

@@

\twocolumn[
\sysmltitle{Guild AI: what to put?}

@@

% You can specify symbols, otherwise they are numbered in order.
% Ideally, you should not use this facility. Affiliations will be numbered
% in order of appearance and this is the preferred way.
\sysmlsetsymbol{equal}{*}

\begin{sysmlauthorlist}
\sysmlauthor{Garrett Smith}{equal,gai}
\end{sysmlauthorlist}

\sysmlaffiliation{gai}{Guild AI, Chicago, Illinois, USA}

@@

\sysmlkeywords{Machine Learning, SysML, Guild AI}

@@

\begin{abstract}
  TODO - Tight summary of Guild, its value proposition wrt
  accelerating research, and the proposed demo audience experience.
\end{abstract}
]

@@

%\printAffiliationsAndNotice{}  % leave blank if no need to mention equal contribution
\printAffiliationsAndNotice{\sysmlEqualContribution} % otherwise use the standard text.

\section{Guild AI}

TODO: What is Guild AI? What is the benefit?

Benefits:

- Trivializes reproducibility - saves time, let's researcher focus on
novel work

- Baselines are easy to integrate into a project

\subsection{Users}

TODO: Who is Guild intended for?

@@

TODO: What other projects

- Model DB

- Pip

- ???

@@

TODO: What makes Guild unique?

- Simplicity and ease of use (talk about YAML file)

- Immediate benefit to users (value of reproducibility extends to
original researcher as well as surveyors)

- Consistent with traditional methods of artifact reuse (packages,
file system based, etc.)

\section{Demonstration}

TODO: Describe the format, approach, etc.

\subsection{Audience Experience}

TODO: What will the audience experience? What will they see? How will
they feel inside?

\subsection{Equipment}

TODO: What sort of equipment is needed?
